Ubercalibration Simulations \\(Euclid)

### 2011-01-31 Documentation/ubercal.tex
euclid: document update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17373 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\documentclass[12pt,a4paper,twoside]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{color}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{upgreek}
\usepackage{subfigure}

@@

\section{Introduction}
These simulations...

\section{Sky Catalog}
We create a fake sky, based on realistic star densities, between a magnitude of 19 to 22.5. The upper and lower limits are defined based the saturation and 10$\upsigma$ limits for a single Euclid exposure. Star's are generated with random coordinates (within the sky region being investigated) and with a random magnitude according to the following powerlaw:
\begin{equation}
\frac{1}{N} \frac{dN}{dm} = a + b m
\end{equation}

@@

\begin{figure}[ht]
\centering
\subfigure[A histogram of the generated stars' magnitudes.]{
\includegraphics[width=0.45\textwidth]{source_mag_histogram.png}
\label{fig:source_magnitude}
}
\subfigure[A small subset of the simulated sky catalog shown on the fake sky.]{
\includegraphics[width=0.45\textwidth]{gods_sky.png}
\label{fig:gods_sky}
}
\label{fig:catalog}
\end{figure}

The random number generator is given the same seed, so that the same simulated sky is produced each time the routine is called.

\section{Single Exposure}
With a given pointing and orientation, we transform the sky catalog into focal plane coordinates and find all the stars falling within the field-of-view. We do not simulate images, instead we trivially convert the actual star fluxes into measured fluxes.

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{flux_uncertainty_variance.png}
\end{center}
\caption{The flux uncertainty used as a function of star magnitude.\label{fig:flux_uncertainty}}
\end{figure}
\subsection{Catalog to Measured Flux Conversion}
The measured flux from a star depends on its focal plane position ($\vec{x_i}$), God's flat-field (with parameters $\vec{q}$) and the star's true flux ($s$).

@@

@@

XX Our model is

@@

\appendix
\section{Random Magnitude Generation}
XX DAVID: I have included this so that I remember how the formula was derived...

@@

\begin{equation}
m = \frac{1}{b} \log{(\left[ 10^{b m_{max}} - 10^{b m_{min}} \right] r + 10^{b m_{min}})}
\end{equation}
\noindent{}where r is a random number in the range [0.0, 1.0).

\end{document}

>>>>

\documentclass[12pt,a4paper,twoside]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{color}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{upgreek}
\usepackage{subfigure}
\usepackage{verbatim}

@@

\section{Sky Catalog}
We create a fake sky, based on realistic star densities, between a magnitude of 19 to 22.5. The upper and lower limits are defined based the saturation and 10$\upsigma$ limits for a single Euclid exposure. Stars are generated with random coordinates (within the sky region being investigated) and with a random magnitude according to the following powerlaw:
\begin{equation}
\frac{1}{N} \frac{dN}{dm} = a + b m
\end{equation}

@@

\begin{figure}[ht]
\centering
\subfigure[A histogram of the generated stars' magnitudes.]{
\includegraphics[width=0.45\textwidth]{source_mag_histogram.png}
\label{fig:source_magnitude}
}
\subfigure[A small subset of the simulated stars in the fake sky.]{
\includegraphics[width=0.45\textwidth]{gods_sky.png}
\label{fig:gods_sky}
}
\label{fig:catalog}
\end{figure}

The random number generator is given the same seed, so that the same simulated sky is produced each time the simulation is run.

\section{Single Exposure}
With a given pointing and orientation, we transform the sky catalog into focal plane coordinates and find all the stars falling within the instrument's field-of-view. The default field-of-view size is the same as the Euclid NISP instrument's. We do not simulate images, instead we trivially convert the actual star fluxes into measured fluxes.

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{flux_uncertainty_variance.png}
\end{center}
\caption{The flux uncertainty used as a function of star magnitude.\label{fig:flux_uncertainty}}
\end{figure}

\subsection{Catalog to Measured Flux Conversion}
The measured flux from a star depends on its focal plane position ($\vec{x_i}$), God's flat-field (with parameters $\vec{q}$) and the star's true flux ($s$).

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{camera.png}
\end{center}
\caption{A single exposure of the sky with the pointing coordinates and the field orientation given in the title. Left: Plot of the entire fake sky, with the instrument's field-of-view shown in black and the measured sources highlighted in red. Right: The distribution of the highlighted sources on the instrument's focal plane. \label{fig:camera}}
\end{figure}

@@

Our model is

@@

\section{Initial Results}
\subsection{Surveys}
To test the code we considered two extreme survey examples. The first, called Survey A, consists of a regular grid of pointings over the sky with a 2.5\% overlap between neighboring fields. Nine passes are performed over the sky and each time the grids are aligned exactly. The second survey, Survey D, has the same number of fields as Survey A, but the pointing and orientation of these fields are random. These two surveys can be seen in Fig. \ref{fig:surveys}.

\begin{figure}[ht]
\centering
\subfigure{
\includegraphics[width=0.47\textwidth]{A.pdf}
}
\subfigure{
\includegraphics[width=0.47\textwidth]{D.pdf}
}
\caption{Left: The A Survey has regular pointings over the sky with a 2.5\% overlap between adjacent fields. Nine passes are performed over the sky and each time the grids are aligned exactly. Right: The D Survey has the same number of pointings as the A Survey, but the pointings and orientations are random.\label{fig:surveys}}
\end{figure}

\subsection{Ubercalibration Performance}

The difference in ubercalibration method's ability to fit the flat field and the star magnitudes for the two different surveys is striking. With the random survey (D) we can get very close to the actual flat field in a small number of ubercalibration iterations (see Fig \ref{fig:D}). The well ordered survey (A), on the other hand, does not converge to the correct flat field even with 1,000 ubercal iterations. Even when we start with almost the correct flat-field parameters (those found with the D survey - shown in Fig. \ref{fig:D}), the ubercalibration procedure does not converge. In fact, with each iteration solution drifts away from the actual flat-field. Fig \ref{fig:A} shows the fitted flat-field from Survey A after 1,000 with the initial parameters shown in Fig \ref{fig:D}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{D_15_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field found with Survey D. \label{fig:D}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{A_1000_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field calculated found with Survey A. The ubercalibration procedure was started with almost exact flat field parameters (those found from Survey D - see Fig. \ref{fig:D}) and has drifted away from the actual solution. \label{fig:A}}
\end{figure}

\appendix
\begin{comment}
\section{Random Magnitude Generation}
XX DAVID: I have included this so that I remember how the formula was derived...

@@

\begin{equation}
m = \frac{1}{b} \log{(\left[ 10^{b m_{max}} - 10^{b m_{min}} \right] r + 10^{b m_{min}})}
\end{equation}
\noindent{}where r is a random number in the range [0.0, 1.0).
\end{comment}
\end{document}
